---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ecocbo)
```

# ecocbo

<!-- badges: start -->
 [![R-CMD-check](https://github.com/arturoSP/ecocbo/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/arturoSP/ecocbo/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of ecocbo is to help scientists to calculate an optimum sampling effort for community ecology projects, based on data from a pilot study only. This package is based on the principles developed on [SSP](https://github.com/edlinguerra/SSP), an R package that simulates ecological communities by extracting and using parameters that control the simulation. The simulated communities are then compared with PERMANOVA, to estimate its componrnts of variation and consequently the optimal sampling effort depending on wether the relevant issue is an economic budget or a required level of precision. 

## Installation

You can install the development version of ecocbo from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("edlinguerra/ecocbo")
```

## Example

This is a basic example which shows you how to use the different functions in the package:

```{r eval=FALSE}
# Prepare the data.
# Load data and adjust it.
data(epiDat)

epiH0 <- epiDat
epiH0[,"site"] <- as.factor("T0")
epiHa <- epiDat
epiHa[,"site"] <- as.factor(epiHa[,"site"])

# Calculate simulation parameters.
parH0 <- SSP::assempar(data = epiH0, type = "counts", Sest.method = "average")
parHa <- SSP::assempar(data = epiHa, type = "counts", Sest.method = "average")

# Simulation.
simH0Dat <- SSP::simdata(parH0, cases = 3, N = 1000, sites = 1)
simHaDat <- SSP::simdata(parHa, cases = 3, N = 100, sites = 10)

```


```{r}
# Calculate statistical power.
betaResult <- sim_beta(simH0 = simH0Dat, simHa = simHaDat, 
                       n = 10, m = 3, k = 20, alpha = 0.05)
betaResult

```


```{r}
# Plot the power progression as sampling increases.
plot_power(data = betaResult, n = NULL, m = 3, method = "power")

```


```{r}
#Calculate components of variation.
compVar <- scompvar(data = betaResult)
compVar

```


```{r}
# Determine optimal sampling effort
cboResult <- sim_cbo(comp.var = compVar, ct = 20000, ck = 100, cj = 2500)
cboResult

```
